---
title: ''
author: ''
date: ''
output: 
  html_document:
    df_print: paged
    keep_md: true
  word_document:
    reference_docx: style.1.docx
---

```{r setup, include = FALSE}

  knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
                        fig.width  = 6 * 1.67, fig.height = 6)

  load("data.RData")
  
  library(tidyverse)
  library(viridis)
  library(flextable)
  library(sjPlot)
  library(sjmisc)
  library(rms)
  library(Hmisc)
  

```

# Overview

**Original request:**

*I have attached a rough draft of the methodology for the Happy Talk study with some research questions which hopefully will help make sense of the data. *

*I have put in the following questions and highlighted the outcomes. In some places there is an option to use just one outcome or to add in another. I don’t know if our numbers allow for a second outcome in some of these places, so statistically you might not advise it.*

*Research questions – Following completion of the Happy Talk programme*

*- Based on the PLS-5 composite scores, did the receptive and or expressive language of children from disadvantaged areas improve, when compared to a control group over the same period of time? PLS-5 outcomes*

*- Using the PedsQL and CHU9D, did children’s quality of life improve compared to a control group of children from similar backgrounds, who did not complete the programme? PEDsQL total score and then composites?*

*- Using the Detailed Code of Responsiveness, did parents increase their level of responsiveness when playing with their children, compared to a control group of parents who did not complete the programme? Rate per minute outcome and total number as a possible second for PedsQL* 



*- When measured using the Communication Supporting Classroom Observation Tool (CSCOT) has the oral language environment of preschools increased when compared to control preschools in similar areas catering for a similar demographic. CSCOT score outcome*
**Note: n = 4. There really isn't an analysis possible here.** 


**Note: The following two questions can not be usefully answered with these data.** 

*- Are attendance rates associated with differential intervention outcome effects?*

*- Examining the influences of moderating or mediating influences on children’s intervention effects.*


# PLS-5

This section answers the following: *Based on the PLS-5 composite scores, did the receptive and or expressive language of children from disadvantaged areas improve, when compared to a control group over the same period of time? PLS-5 outcomes*

## PLS-5-AC

Figure. Individual level change in PLS-5-AC by study arm. 

```{r plot_code}

  dist_plot <- function(var, lab){
    ggplot(long, aes_string(y = var, x = "time", group = "code", color = "arm", 
                            fill = "arm")) +
      geom_line(alpha = 0.5, color = "white", linetype = "dotted") +
      geom_smooth(aes(group = arm), 
                 method = "lm", se = FALSE, color = "white", size = 2) +
      geom_point(position = position_jitter(width = 0.01), size = 3, 
                 alpha = 0.6, 
                 color = "white") +
      geom_boxplot(data = filter(long, time == "Pre"), 
                   aes(group = time), alpha = 0.8, color = "white", 
                   outlier.shape = NA, width = 0.2, 
                   position = position_nudge(x = -0.25)) +
      geom_boxplot(data = filter(long, time == "Post"), 
                     aes(group = time), alpha = 0.8, color = "white", 
                     outlier.shape = NA, width = 0.2, 
                     position = position_nudge(x = 0.25)) +
    # geom_flat_violin(aes(group = time), alpha = 0.2, 
    #                  adjust = 1, trim = TRUE, color = NA) + 
      geom_violin(data = filter(long, time == "Post"),
                  aes(group = time), alpha = 0.3, 
                  adjust = 1, trim = TRUE, color = NA, 
                  scale = "count", width = 0.3,
                  position = position_nudge(x = 0.25)) + 
      geom_violin(data = filter(long, time == "Pre"),
                  aes(group = time), alpha = 0.3, 
                  adjust = 1, trim = TRUE, color = NA, 
                  scale = "count", width = 0.3,
                  position = position_nudge(x = -0.25)) + 
      facet_wrap(~arm) +
      scale_fill_viridis(guide = FALSE, discrete = TRUE, begin = 0.5, 
                         end = 0.8, option = "inferno", direction = -1) +
      scale_color_viridis(guide = FALSE, discrete = TRUE, begin = 0.5, 
                          end = 0.8, option = "inferno", direction = -1) +
      xlab("") +
      ylab(lab) +
      theme(panel.background = element_rect(fill = "grey20"), 
            panel.grid = element_blank(), 
            strip.background = element_rect(fill = "grey20"), 
            strip.text = element_text(color = "white")) %>%
    return()
  }

```



```{r pls5_ac_line}

  dist_plot("pls5_ac", "PLS-5 AC")

```


Table. Estimated effect of the intervention on PLS-5-AC scores.
```{r pls5_ac_models}

# baseline adjusted
  pls5_ac_m1 <- lm(pls5_ac_post ~ arm + center(pls5_ac_pre), 
                   data = data) 
# Plus age and slt
  pls5_ac_m2 <- lm(pls5_ac_post ~ arm + center(pls5_ac_pre) + 
                     center(age_in_months) + previous_slt_involvement, 
                   data = data) 
 
  tab_model(
    list(pls5_ac_m1, pls5_ac_m2),
    dv.labels = c("Baseline adjusted", "Covariate adjusted"),
    pred.labels = c("Intercept", "Active (vs Control)", "Baseline PLS-5-AC", 
                    "Age (m)", "Previous SLT (vs None)")
    )
  

```

## PLS-5-ES

Figure. Individual level change in PLS-5-ES by study arm. 
```{r pls5_es_line}

  dist_plot("pls5_es", "PLS-5 ES")

```

Table. Estimated effect of the intervention on PLS-5-ES scores.
```{r pls5_es_models}

# baseline adjusted
  pls5_es_m1 <- lm(pls5_es_post ~ arm + center(pls5_es_pre), 
                   data = data) 
# Plus age and slt
  pls5_es_m2 <- lm(pls5_es_post ~ arm + center(pls5_es_pre) + 
                     center(age_in_months) + previous_slt_involvement, 
                   data = data) 
 
  tab_model(
    list(pls5_es_m1, pls5_es_m2),
    dv.labels = c("Baseline adjusted", "Covariate adjusted"),
    pred.labels = c("Intercept", "Active (vs Control)", "Baseline PLS-5-ES", 
                    "Age (m)", "Previous SLT (vs None)")
    )
  

```


## PLS-5-Total

Figure. Individual level change in PLS-5-Total by study arm. 
```{r pls5_to_line}

  dist_plot("pls5_to", "PLS-5-Total")

```


Table. Estimated effect of the intervention on PLS-5-Total scores.
```{r pls5_to_models}

# baseline adjusted
  pls5_to_m1 <- lm(pls5_to_post ~ arm + center(pls5_to_pre), 
                   data = data) 
# Plus age and slt
  pls5_to_m2 <- lm(pls5_to_post ~ arm + center(pls5_to_pre) + 
                     center(age_in_months) + previous_slt_involvement, 
                   data = data) 
 
  tab_model(
    list(pls5_to_m1, pls5_to_m2),
    dv.labels = c("Baseline adjusted", "Covariate adjusted"),
    pred.labels = c("Intercept", "Active (vs Control)", "Baseline PLS-5-Total", 
                    "Age (m)", "Previous SLT (vs None)")
    )
  
```

# QOL

This section answers the following: 

*- Using the PedsQL and CHU9D, did children’s quality of life improve compared to a control group of children from similar backgrounds, who did not complete the programme? PEDsQL total score and then composites?*

## PEDsQL-Psychosocial

Figure. Individual level change in PEDsQL-Psychosocial scores by study arm. 
```{r pedsql_ps_line}

  dist_plot("peds_psychosocial", "PEDsQL-Psychosocial")

```

Table. Estimated effect of the intervention on PEDsQL-Psychosocial scores.
```{r pedsql_ps_models}

# baseline adjusted
  pedsql_ps_m1 <- orm(post_peds_psychosocial ~ arm +
                        center(peds_psychosocial_pre), 
                    data = data) 

# Plus age and slt
  pedsql_ps_m2 <- orm(post_peds_psychosocial ~ arm +
                        center(peds_psychosocial_pre) + 
                        center(age_in_months) + previous_slt_involvement, 
                    data = data) 

  options(prType="html")
   
  print(pedsql_ps_m1)
 
  print(pedsql_ps_m2)

  
```

## PEDsQL-Physical

Figure. Individual level change in PEDsQL-Physical scores by study arm. 
```{r pedsql_ph_line}

  dist_plot("peds_physical", "PEDsQL-Physical")

```


Table. Estimated effect of the intervention on PEDsQL-Physical scores.
```{r pedsql_ph_models}

# baseline adjusted
  pedsql_ph_m1 <- orm(post_peds_physical ~ arm +
                        center(peds_physical_pre), 
                    data = data) 

# Plus age and slt
  pedsql_ph_m2 <- orm(post_peds_physical ~ arm +
                        center(peds_physical_pre) + 
                        center(age_in_months) + previous_slt_involvement, 
                    data = data) 

  options(prType="html")
   
  print(pedsql_ph_m1)
 
  print(pedsql_ph_m2)

  
```

## PEDsQL-Total

Figure. Individual level change in PEDsQL-Total scores by study arm. 
```{r pedsql_to_line}

  dist_plot("peds_tot", "PEDsQL-Total")

```



Table. Estimated effect of the intervention on PEDsQL-Total scores.
```{r pedsql_to_models}

# baseline adjusted
  pedsql_to_m1 <- orm(post_peds_tot ~ arm +
                        center(peds_tot_pre), 
                    data = data) 

# Plus age and slt
  pedsql_to_m2 <- orm(post_peds_tot ~ arm +
                        center(peds_tot_pre) + 
                        center(age_in_months) + previous_slt_involvement, 
                    data = data) 

  options(prType="html")
   
  print(pedsql_to_m1)
 
  print(pedsql_to_m2)

  
```

## CHU9D-Total

Figure. Individual level change in CHU9D-Total scores by study arm. 
```{r chu9d_to_line}

  dist_plot("chu9d_tot", "CHU9D-Total")

```


Table. Estimated effect of the intervention on PEDsQL-Total scores.
```{r chu9d_to_models}

# baseline adjusted
  chu9d_to_m1 <- orm(post_chu9d_tot ~ arm +
                        center(chu9d_tot_pre), 
                    data = data) 

# Plus age and slt
  chu9d_to_m2 <- orm(post_chu9d_tot ~ arm +
                        center(chu9d_tot_pre) + 
                        center(age_in_months) + previous_slt_involvement, 
                    data = data) 

  options(prType="html")
   
  print(chu9d_to_m1)
 
  print(chu9d_to_m2)

  
```


# Responsiveness

This section answers the following: 

*- Using the Detailed Code of Responsiveness, did parents increase their level of responsiveness when playing with their children, compared to a control group of parents who did not complete the programme? Rate per minute outcome and total number as a possible second for PedsQL*

## Expansions

Figure. Individual level change in Expansions by study arm. 
```{r exp_line}

  dist_plot("res_exp", "Expansions")

```


Table. Estimated effect of the intervention on Expansions.
```{r expansion_models}

# baseline adjusted
  res_exp_m1 <- lm(res_exp_post ~ arm + center(res_exp_pre), 
                    data = data) 

# Plus age and slt
  res_exp_m2 <- lm(res_exp_post ~ arm + center(res_exp_pre) + 
                     center(age_in_months) + previous_slt_involvement, 
                   data = data) 

  tab_model(
    list(res_exp_m1, res_exp_m2),
    dv.labels = c("Baseline adjusted", "Covariate adjusted"),
    pred.labels = c("Intercept", "Active (vs Control)", "Baseline Expansions", 
                    "Age (m)", "Previous SLT (vs None)")
    )

  
```


## Imitations

Figure. Individual level change in Imitations by study arm. 
```{r im_line}

  dist_plot("res_im", "Imitations")

```


Table. Estimated effect of the intervention on Imitations.
```{r imitations_models}

# baseline adjusted
  res_im_m1 <- lm(res_im_post ~ arm + center(res_im_pre), 
                    data = data) 

# Plus age and slt
  res_im_m2 <- lm(res_im_post ~ arm + center(res_im_pre) + 
                     center(age_in_months) + previous_slt_involvement, 
                   data = data) 

  tab_model(
    list(res_im_m1, res_im_m2),
    dv.labels = c("Baseline adjusted", "Covariate adjusted"),
    pred.labels = c("Intercept", "Active (vs Control)", "Baseline Imitations", 
                    "Age (m)", "Previous SLT (vs None)")
    )

  
```


## Responsive Questions

Figure. Individual level change in Responsive Questions by study arm. 
```{r responsive_line}

  dist_plot("res_res", "Responsive Questions")

```


Table. Estimated effect of the intervention on Responsive Questions.
```{r responsive_models}

# baseline adjusted
  res_res_m1 <- lm(res_res_post ~ arm + center(res_res_pre), 
                    data = data) 

# Plus age and slt
  res_res_m2 <- lm(res_res_post ~ arm + center(res_res_pre) + 
                     center(age_in_months) + previous_slt_involvement, 
                   data = data) 

  tab_model(
    list(res_res_m1, res_res_m2),
    dv.labels = c("Baseline adjusted", "Covariate adjusted"),
    pred.labels = c("Intercept", "Active (vs Control)", 
                    "Baseline Responsive Questions", 
                    "Age (m)", "Previous SLT (vs None)")
    )

  
```

## Labels

Figure. Individual level change in Labels by study arm. 
```{r label_line}

  dist_plot("res_label", "Labels")

```


Table. Estimated effect of the intervention on Labels.
```{r labels_models}

# baseline adjusted
  res_label_m1 <- lm(res_label_post ~ arm + center(res_label_pre), 
                    data = data) 

# Plus age and slt
  res_label_m2 <- lm(res_label_post ~ arm + center(res_label_pre) + 
                     center(age_in_months) + previous_slt_involvement, 
                   data = data) 

  tab_model(
    list(res_label_m1, res_label_m2),
    dv.labels = c("Baseline adjusted", "Covariate adjusted"),
    pred.labels = c("Intercept", "Active (vs Control)", "Baseline Labels", 
                    "Age (m)", "Previous SLT (vs None)")
    )

  
```

## Responsiveness Total

Figure. Individual level change in Responsiveness Total by study arm. 
```{r total_line}

  dist_plot("res_total", "Responsiveness Total")

```


Table. Estimated effect of the intervention on Responsiveness Total.
```{r total_models}

# baseline adjusted
  res_total_m1 <- lm(res_total_post ~ arm + center(res_total_pre), 
                    data = data) 

# Plus age and slt
  res_total_m2 <- lm(res_total_post ~ arm + center(res_total_pre) + 
                     center(age_in_months) + previous_slt_involvement, 
                   data = data) 

  tab_model(
    list(res_total_m1, res_total_m2),
    dv.labels = c("Baseline adjusted", "Covariate adjusted"),
    pred.labels = c("Intercept", "Active (vs Control)", 
                    "Baseline Responsiveness Total", 
                    "Age (m)", "Previous SLT (vs None)")
    )

  
```

## Responsiveness Rate

Figure. Individual level change in Responsiveness Rate by study arm. 
```{r rate_line}

  dist_plot("res_rate", "Responsiveness Rate")

```


Table. Estimated effect of the intervention on Responsiveness Total.
```{r rate_models}

# baseline adjusted
  res_rate_m1 <- lm(res_rate_post ~ arm + center(res_rate_pre), 
                    data = data) 

# Plus age and slt
  res_rate_m2 <- lm(res_rate_post ~ arm + center(res_rate_pre) + 
                     center(age_in_months) + previous_slt_involvement, 
                   data = data) 

  tab_model(
    list(res_rate_m1, res_rate_m2),
    dv.labels = c("Baseline adjusted", "Covariate adjusted"),
    pred.labels = c("Intercept", "Active (vs Control)", 
                    "Baseline Responsiveness Total", 
                    "Age (m)", "Previous SLT (vs None)")
    )
  
```


## PaRRiS

Figure. Individual level change in PaRRiS by study arm. 
```{r parris_line}

  dist_plot("parris", "PaRRiS")

```


Table. Estimated effect of the intervention on PaRRiS.
```{r parris_models}

# baseline adjusted
  parris_m1 <- lm(parris_post ~ arm + center(parris_pre), 
                    data = data) 

# Plus age and slt
  parris_m2 <- lm(parris_post ~ arm + center(parris_pre) + 
                     center(age_in_months) + previous_slt_involvement, 
                   data = data) 

  tab_model(
    list(parris_m1, parris_m2),
    dv.labels = c("Baseline adjusted", "Covariate adjusted"),
    pred.labels = c("Intercept", "Active (vs Control)", 
                    "Baseline PaRRiS", 
                    "Age (m)", "Previous SLT (vs None)")
    )

  
```


```{r publication_qualityplots}

# ggsave("XXX", device = "pdf", dpi = 600)

```

```{r if_html, eval = knitr::opts_knit$get("rmarkdown.pandoc.to") == "html"}

# use flextable (works for MD files on github too)

```

```{r if_word, eval = !knitr::opts_knit$get("rmarkdown.pandoc.to") == "html"}

# use kable

```

```{r code_book}

  print(summarytools::dfSummary(data), method = "render")

```

```{r sysinfo}

  DescTools::SysInfo()

```


