---
title: ''
author: ''
date: ''
output: 
  html_document:
    df_print: paged
    keep_md: true
  word_document:
    reference_docx: style.1.docx
---

```{r setup, include = FALSE}

  knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
                        fig.width  = 6 * 1.67, fig.height = 6)

  load("data.RData")
  
  library(tidyverse)
  library(viridis)
  library(flextable)
  library(sjPlot)
  library(sjmisc)

```

**Original request:**

I have attached a rough draft of the methodology for the Happy Talk study with some research questions which hopefully will help make sense of the data. 
I have put in the following questions and highlighted the outcomes. In some places there is an option to use just one outcome or to add in another. I don’t know if our numbers allow for a second outcome in some of these places, so statistically you might not advise it. 

Research questions – Following completion of the Happy Talk programme

- Based on the PLS-5 composite scores, did the receptive and or expressive language of children from disadvantaged areas improve, when compared to a control group over the same period of time? PLS-5 outcomes

- Using the PedsQL and CHU9D, did children’s quality of life improve compared to a control group of children from similar backgrounds, who did not complete the programme? PEDsQL total score and then composites?

- Using the Detailed Code of Responsiveness, did parents increase their level of responsiveness when playing with their children, compared to a control group of parents who did not complete the programme? Rate per minute outcome and total number as a possible second for PedsQL

- When measured using the Communication Supporting Classroom Observation Tool (CSCOT) has the oral language environment of preschools increased when compared to control preschools in similar areas catering for a similar demographic. CSCOT score outcome
```{r}



```

There are also some other questions for example whether attendance rates (which I think were quite poor) had any impact on the child language or parent video outcomes – but again I know that some of this will depend on what is possible given our numbers.

# PLS-5

- Based on the PLS-5 composite scores, did the receptive and or expressive language of children from disadvantaged areas improve, when compared to a control group over the same period of time? PLS-5 outcomes

## PLS-5-AC

```{r pls5_ac_line}

  ggplot(long, aes(y = pls5_ac, x = time, group = code, color = arm)) +
    geom_line() +
    facet_wrap(~arm) +
    scale_color_viridis("", discrete = TRUE, end = 0.5) +
    xlab("") +
    ylab("PLS-5 AC")

```

```{r pls5_ac_density}

  ggplot(long, aes(x = pls5_ac, fill = time, color = time)) +
    geom_density(alpha = 0.5, color = "black") +
    geom_rug(size = 2, alpha = 0.5) +
    facet_wrap(~arm, nrow = 2) +
    scale_fill_viridis("", discrete = TRUE, end = 0.5) +
    scale_color_viridis("", discrete = TRUE, end = 0.5) +
    ylab("Density") +
    xlab("PLS-5 AC")

```
```{r pls5_ac_models}

# baseline adjusted
  pls5_ac_m1 <- lm(pls5_ac_post ~ arm + center(pls5_ac_pre), 
                   data = data) 
# Plus age and slt
  pls5_ac_m2 <- lm(pls5_ac_post ~ arm + center(pls5_ac_pre) + 
                     center(age_in_months) + previous_slt_involvement, 
                   data = data) 
 
  tab_model(
    list(pls5_ac_m1, pls5_ac_m2),
    dv.labels = c("Basline adjusted", "Covariate adjusted"),
    pred.labels = c("Intercept", "Active (vs Control)", "Baseline PLS-5-AC", 
                    "Age (m)", "Previous SLT (vs None)")
    )
  

```

## PLS-5-ES

```{r pls5_es_line}

  ggplot(long, aes(y = pls5_es, x = time, group = code, color = arm)) +
    geom_line() +
    facet_wrap(~arm) +
    scale_color_viridis("", discrete = TRUE, end = 0.5) +
    xlab("") +
    ylab("PLS-5 ES")

```

```{r pls5_es_density}

  ggplot(long, aes(x = pls5_es, fill = time, color = time)) +
    geom_density(alpha = 0.5, color = "black") +
    geom_rug(size = 2, alpha = 0.5) +
    facet_wrap(~arm, nrow = 2) +
    scale_fill_viridis("", discrete = TRUE, end = 0.5) +
    scale_color_viridis("", discrete = TRUE, end = 0.5) +
    ylab("Density") +
    xlab("PLS-5 ES")

```

```{r pls5_es_models}

# baseline adjusted
  pls5_es_m1 <- lm(pls5_es_post ~ arm + center(pls5_es_pre), 
                   data = data) 
# Plus age and slt
  pls5_es_m2 <- lm(pls5_es_post ~ arm + center(pls5_es_pre) + 
                     center(age_in_months) + previous_slt_involvement, 
                   data = data) 
 
  tab_model(
    list(pls5_es_m1, pls5_es_m2),
    dv.labels = c("Basline adjusted", "Covariate adjusted"),
    pred.labels = c("Intercept", "Active (vs Control)", "Baseline PLS-5-ES", 
                    "Age (m)", "Previous SLT (vs None)")
    )
  

```


## PLS-5-Total

```{r pls5_to_line}

  ggplot(long, aes(y = pls5_to, x = time, group = code, color = arm)) +
    geom_line() +
    facet_wrap(~arm) +
    scale_color_viridis("", discrete = TRUE, end = 0.5) +
    xlab("") +
    ylab("PLS-5 ES")

```

```{r pls5_to_density}

  ggplot(long, aes(x = pls5_to, fill = time, color = time)) +
    geom_density(alpha = 0.5, color = "black") +
    geom_rug(size = 2, alpha = 0.5) +
    facet_wrap(~arm, nrow = 2) +
    scale_fill_viridis("", discrete = TRUE, end = 0.5) +
    scale_color_viridis("", discrete = TRUE, end = 0.5) +
    ylab("Density") +
    xlab("PLS-5 ES")

```

```{r pls5_to_models}

# baseline adjusted
  pls5_to_m1 <- lm(pls5_to_post ~ arm + center(pls5_to_pre), 
                   data = data) 
# Plus age and slt
  pls5_to_m2 <- lm(pls5_to_post ~ arm + center(pls5_to_pre) + 
                     center(age_in_months) + previous_slt_involvement, 
                   data = data) 
 
  tab_model(
    list(pls5_to_m1, pls5_to_m2),
    dv.labels = c("Basline adjusted", "Covariate adjusted"),
    pred.labels = c("Intercept", "Active (vs Control)", "Baseline PLS-5-ES", 
                    "Age (m)", "Previous SLT (vs None)")
    )
  
```


```{r publication_qualityplots}

# ggsave("XXX", device = "pdf", dpi = 600)

```

```{r if_html, eval = knitr::opts_knit$get("rmarkdown.pandoc.to") == "html"}

# use flextable (works for MD files on github too)

```

```{r if_word, eval = !knitr::opts_knit$get("rmarkdown.pandoc.to") == "html"}

# use kable

```

```{r code_book}

  print(summarytools::dfSummary(data), method = "render")

```

```{r sysinfo}

  DescTools::SysInfo()

```


